---
title: "tidytuesdaywk5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r read_data, include=FALSE}
library(here)
library(tidyverse)
library(tidytuesdayR)
library(janitor)
library(ggimage)
library(ggpattern)

breed_traits <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-02-01/breed_traits.csv') %>%
  clean_names()
trait_description <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-02-01/trait_description.csv')
breed_rank_all <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-02-01/breed_rank.csv') %>%
  clean_names()
```

There are 195 breeds of dog in both the breed traits and rank table. 
I want to join these tables together. 

```{r}
nrow(breed_traits)
nrow(breed_rank_all)
```

However, there's a problem with the whitespace that means that the names of breeds do not match between tables. 

```{r}
gsub("\\s", "", breed_traits$breed[[1]])
gsub("\\s", "", breed_rank_all$breed[[1]])

# space is 160, non-breaking space,
# rather than 32, regular space
print(utf8ToInt(breed_traits$breed[[1]]))
print(utf8ToInt(breed_rank_all$breed[[1]]))

# let's change it to a regular space...
breed_traits$breed <- gsub("\u00A0", " ", breed_traits$breed)
```

Now all but one of the breeds match between the two tables. This breed has an apostrophe instead of a space in one of the tables. I remove the apostrophe. 

```{r}
setdiff(breed_traits$breed, breed_rank_all$breed)

breed_traits$breed[str_detect(breed_traits$breed, "Cirnechi")]
breed_rank_all$breed[str_detect(breed_rank_all$breed, "Cirnechi")]

breed_traits$breed <- gsub("’", " ", breed_traits$breed)

setdiff(breed_traits$breed, breed_rank_all$breed)
```

Now the two tables can be joined using breed as a key. 
For some reason the Plott hound doesn't have a valid coat type, so that record is dropped. 

```{r}
breeds <- breed_traits %>%
  left_join(breed_rank_all, by = "breed") %>%
  filter(breed != "Plott Hounds")
nrow(breeds)
head(breeds)
```

Next, for each coat type I identify the most commonly registered dog with that type of coat in 2020.
This is to ideally provide an example dog that the viewer will be able to recognise. 

```{r}
examples <- breeds %>%
  # get the dog people will be most familiar with
  arrange(x2020_rank) %>%
  group_by(coat_type) %>%
  # first row will have the highest rated dog for that coat type
  filter(row_number() == 1) %>%
  select(breed, coat_type, image) %>%
  mutate(texture = here("TidyTuesday", paste0(tolower(coat_type), "_coat.png")))
examples
```

Next I create a simple bar chart demonstrating the numbers of AKC dog breeds with each coat type. 

```{r}
breeds %>%
  filter(coat_type != "Plott Hounds") %>%
  group_by(coat_type) %>%
  summarise(count = n()) %>%
  arrange(count) %>%
  left_join(examples, by = "coat_type") %>%
  ungroup() %>%
ggplot(aes(x = count, y = fct_reorder(coat_type, count))) +
  #geom_col(colour = "black", fill = "lightgrey") +
  geom_col_pattern(
    aes(pattern_filename = I(texture)),
    pattern          = 'image', 
    pattern_type     = 'expand',
    colour           = 'black'
  ) +
  geom_text(aes(label = paste("e.g. ", breed), x = count + 10), hjust = 0, size = 2.5) +
  geom_image(aes(y = seq(1, nrow(examples)), x = count + 5, 
                 image = fct_reorder(image, count)),
             size = 0.07, asp = 1.6) +
  theme_classic() +
  theme(axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_line(size = 0.1, color = "lightgrey")) +
  ggtitle("Frequency of Coat Types in AKC Dog Breeds") +
  xlab("# of Breeds") + 
  ylab("") +
  xlim(0, 95)
ggsave(here("TidyTuesday", "akc_coat_types.png"),
       height = 3.7,
       width = 5.9)
```